HiC Contact Map Comaprison Using Graphlet Approach

### 2018-04-26 papers/manuscripts/manuscript.tex
new statistical analysis: Manova

In this study, we plan to find dissimilarities between normal cells and cancerous cells,
through investigating HiC contact maps. 
We suspect that there are systematic differences between how chromosomes are structured
between normal cells and cancerous cells.
we 
Ideally, it is desirable to compare 3D structures of 
cell in order to make such comparisons.
However, the main challenge that we face is that 
3D structure of a cell is not readily available. Based on
\cite{adhikari2016chromosome3d}, fluorescence in situ hybridizaiton
(FISH) is used for investigating 3D configuration of chromosomes.
However, this method can only be used locally and cannot map
the whole structure of the chromosomes.
In orther to find dissimilarities in the 3D structure of 
chromosomes, we used HiC dataset.
The HiC method, which was developed by \cite{lieberman2009comprehensive}, captures interactions between 
chromosomal fragments in kilobase resolution. Based on HiC data, an
\textit{interaction frequency (IF) } matrix can be developed between \textit{loci} at a desired resolution.
A cell $IF_{ij}$ in an interaction frequency matrix captures the number of interaction detected
in HiC dataset between locus $i$ and locus $j$ in the genome.
An interaction matrix can be used to develop both inter- and intra-chromosomal interaction matrices.
We believe differences in interaction matrices can be found between normal cells and cancerous ones.

@@

\cite{milenkoviae2008uncovering} 
 provide heuristics to compare two nodes based on some feature
(or signature) vectors, which is a 73-dimensional vector
$\mathbf{s}^T
= [s_0, s_2, ..., s_{72}]$ where $s_i$ denotes the number of nodes in
the network that are part of an orbit $i$. \\
They concluded that proteins with similar surroundings perform
similar functions.

@@

\subsection{Statistical Analysis}
In order to quntify our results, we have conducted one-sided
t-test in order to test whether the average pair-wise
MIC values across orbits are significantly different from 
each other and also from 1. 
however, pairwise correlation of cancer cells for
orbits in $\mathbb{Q}$ were significantly higher than
correlation between cancer cells and normal cells.
The results
for this test showed that all values are significantly less than 1;
A summary of can be found in
table \ref{tab:t-test}.
We have also conducted t-tests for 
pair-wise orbit signature vector distances across contact maps.
Please refer to supplementary material for result of the 
full list of t-test results.
\begin{table}[H]
    \resizebox{.5\textwidth}{!}{
\begin{tabular}{c|ccccccc}
    & $MIC_{\scriptscriptstyle MIT-CALL4}$ & $MIC_{\scriptscriptstyle MIT-RL}$ & $MIC_{\scriptscriptstyle MIT-ALL}$ & $MIC_{\scriptscriptstyle ALL-CALL4}$ & $MIC_{\scriptscriptstyle CAll4-RL}$ & $MIC_{\scriptscriptstyle ALL-RL}$ & 1   \\
              \rule{0pt}{1pt} \\ \hline \rule{0pt}{15pt}
              $MIC_{\scriptscriptstyle MIT-CALL4}$ & -         & $\mathbf{<}$    & $\mathbf{<}$     & $\mathbf{<}$       & $\mathbf{<}$      & $\mathbf{<}$    & $\mathbf{<}$ \\
              \rule{0pt}{20pt}
              $MIC_{\scriptscriptstyle MIT-RL}$    & $\mathbf{>}$       &   -    & $\mathbf{<}$     & $\mathbf{<}$       & $\mathbf{<}$      & $\mathbf{<}$    & $\mathbf{<}$ \\
              \rule{0pt}{20pt}
              $MIC_{\scriptscriptstyle MIT-ALL}$   & $\mathbf{>}$       & $\mathbf{>}$    &    -    & $\mathbf{<}$       & $\mathbf{<}$      & $\mathbf{<}$    & $\mathbf{<}$ \\
              \rule{0pt}{20pt}
              $MIC_{\scriptscriptstyle ALL-CALL4}$ & $\mathbf{>}$       & $\mathbf{>}$    & $\mathbf{>}$     &       -   & $\mathbf{<}$      & $\mathbf{<}$    & $\mathbf{<}$ \\
              \rule{0pt}{20pt}
              $MIC_{\scriptscriptstyle CAll4-RL}$  & $\mathbf{>}$       & $\mathbf{>}$    & $\mathbf{>}$     & $\mathbf{>}$       &    -     & $\mathbf{<}$    & $\mathbf{<}$ \\
              \rule{0pt}{20pt}
              $MIC_{\scriptscriptstyle ALL-RL}$    & $\mathbf{>}$       & $\mathbf{>}$    & $\mathbf{>}$     & $\mathbf{>}$       & $\mathbf{>}$      &   -    & $\mathbf{<}$ \\
              \rule{0pt}{20pt}
              1         & $\mathbf{>}$       & $\mathbf{>}$    & $\mathbf{>}$     & $\mathbf{>}$       & $\mathbf{>}$      & $\mathbf{>}$    &    -
\end{tabular}}
    \caption{
        Result of one-sided t-test 
        comparison of MIC results 
        with significance $\alpha = 0.01$
        for orbits of $\mathbb{Q}$.
        A $<$ sign means that the value in the corresponding
        row is significantly smaller that the value in 
        corresponding column; and a $>$ sign means the opposite.
        An $=$ sign would have meant the values are not 
        statistically different.
        As can be seen, all MIC values are significantly smaller that 1,
        meaning non on them are exactly the same; 
        however, all Cancer vs Cancer MIC values are statistically
        larger than Cancer vs Normal, implying that 
        spatial positioning of loci are more similar between cancer cells.
    }
    \label{tab:t-test}

>>>>

@@

In \cite{milenkoviae2008uncovering} the authors
 provide heuristics to compare two nodes in a graph
based on signature vectors, which are 73-dimensional vectors
$\mathbf{s}^T
= [s_0, s_2, ..., s_{72}]$ where $s_i$ denotes the number of nodes in
the network that are part of an orbit $i$. \\
They concluded that proteins with similar surroundings perform
similar functions.

@@

\subsection{Statistical Analysis}
In order to qauntify our results, we have conducted one-sided
t-test in order to test whether the average pair-wise
MIC values across orbits are significantly different from 
each other and also from 1. 
however, pairwise correlation of cancer cells for
orbits in $\mathbb{Q}$ were significantly higher than
correlation between cancer cells and normal cells.
The results
for this test showed that all values are significantly less than 1;
A summary of can be found in
table \ref{tab:t-test}.
We have also conducted t-tests for 
pair-wise orbit signature vector distances across contact maps.
Please refer to supplementary material for result of the 
full list of t-test results.
\begin{table}[H]
    \resizebox{.5\textwidth}{!}{
\begin{tabular}{c|ccccccc}
    & $MIC_{\scriptscriptstyle MIT-CALL4}$ & $MIC_{\scriptscriptstyle MIT-RL}$ & $MIC_{\scriptscriptstyle MIT-ALL}$ & $MIC_{\scriptscriptstyle ALL-CALL4}$ & $MIC_{\scriptscriptstyle CAll4-RL}$ & $MIC_{\scriptscriptstyle ALL-RL}$ & 1   \\
              \rule{0pt}{1pt} \\ \hline \rule{0pt}{15pt}
              $MIC_{\scriptscriptstyle MIT-CALL4}$ & -         & $\mathbf{<}$    & $\mathbf{<}$     & $\mathbf{<}$       & $\mathbf{<}$      & $\mathbf{<}$    & $\mathbf{<}$ \\
              \rule{0pt}{20pt}
              $MIC_{\scriptscriptstyle MIT-RL}$    & $\mathbf{>}$       &   -    & $\mathbf{<}$     & $\mathbf{<}$       & $\mathbf{<}$      & $\mathbf{<}$    & $\mathbf{<}$ \\
              \rule{0pt}{20pt}
              $MIC_{\scriptscriptstyle MIT-ALL}$   & $\mathbf{>}$       & $\mathbf{>}$    &    -    & $\mathbf{<}$       & $\mathbf{<}$      & $\mathbf{<}$    & $\mathbf{<}$ \\
              \rule{0pt}{20pt}
              $MIC_{\scriptscriptstyle ALL-CALL4}$ & $\mathbf{>}$       & $\mathbf{>}$    & $\mathbf{>}$     &       -   & $\mathbf{<}$      & $\mathbf{<}$    & $\mathbf{<}$ \\
              \rule{0pt}{20pt}
              $MIC_{\scriptscriptstyle CAll4-RL}$  & $\mathbf{>}$       & $\mathbf{>}$    & $\mathbf{>}$     & $\mathbf{>}$       &    -     & $\mathbf{<}$    & $\mathbf{<}$ \\
              \rule{0pt}{20pt}
              $MIC_{\scriptscriptstyle ALL-RL}$    & $\mathbf{>}$       & $\mathbf{>}$    & $\mathbf{>}$     & $\mathbf{>}$       & $\mathbf{>}$      &   -    & $\mathbf{<}$ \\
              \rule{0pt}{20pt}
              1         & $\mathbf{>}$       & $\mathbf{>}$    & $\mathbf{>}$     & $\mathbf{>}$       & $\mathbf{>}$      & $\mathbf{>}$    &    -
\end{tabular}}
    \caption{
        Result of one-sided t-test 
        comparison of MIC results 
        with significance $\alpha = 0.01$
        for orbits of $\mathbb{Q}$.
        A $<$ sign means that the value in the corresponding
        row is significantly smaller that the value in 
        corresponding column; and a $>$ sign means the opposite.
        An $=$ sign would have meant the values are not 
        statistically different.
        As can be seen, all MIC values are significantly smaller that 1,
        meaning non on them are exactly the same; 
        however, all Cancer vs Cancer MIC values are statistically
        larger than Cancer vs Normal, implying that 
        spatial positioning of loci are more similar between cancer cells.
    }
    \label{tab:t-test}
